---
title: Removing Pesky Unicode Zero Width Spaces
author: "Tiernan"
date: '2018-05-28' 
slug: removing-pesky-unicode-zero-width-spaces
draft: false
tags:
- regex
- tips
categories: R
---

```{r global_options, include=FALSE}
options(tidyverse.quiet = TRUE)
knitr::opts_chunk$set(collapse=TRUE, warning=FALSE, comment='##')

```

```{r load_packages, include=FALSE} 
library(stringi)
library(tidyverse)
```

I recently ran into a problem while downloading a dataset: the character strings contained unwanted zero-width spaces (`<U+200B>`).

These characters are unicode characters used to adjust word spacing and line breaks (see this Wikipedia [article](https://en.wikipedia.org/wiki/Zero-width_space) for more information).

Below is a brief exploration of this particular unicode character followed by two ways to remove it:

  + Targeted approach: `str_replace_all("string","\\u200b","")`
  + Generalized approach: `iconv("string", "utf-8", "ascii", sub="")`

### Exploring the unicode zero-width space character: `<U+200B>`
```{r explore_unicode}
a <- "\u200b"

# Print unicode
print(a)

# Print escaped unicode
stri_escape_unicode(a)

# Replace the unicode with an empty string
a_no_unicode <- str_replace(a,"\\u200b","")

# Print the revised string
print(a_no_unicode)

# Print the escaped unicode version of the revised string (just to double-check)
stri_escape_unicode(a_no_unicode)


```

```{r table}
# Compare the results in a table
tibble("print(a)" = a,
       "stri_escape_unicode(a)" = stri_escape_unicode(a),
       "print(a_no_unicode)" = a_no_unicode,
       "stri_escape_unicode(a_no_unicode)" = stri_escape_unicode(a_no_unicode))

```


### Example with data from the wild
```{r kc-agency-example}
csv_url <- "https://data.kingcounty.gov/resource/es38-6nrz.csv"

agencies <- read_csv(csv_url)

# Print data with unicode characters
print(as.data.frame(agencies[1:5,]))

# Targeting the exact text doesn't work
zwsp_pattern_first <- "<U+200B>"

agencies %>% 
  mutate_all(funs(str_replace_all(.,zwsp_pattern_first,""))) %>% 
  slice(1:5) %>% 
  as.data.frame()

# Escaping the unicode does work
zwsp_pattern_second <- "\\u200b"

agencies %>% 
  mutate_all(funs(str_replace_all(.,zwsp_pattern_second,""))) %>% 
  slice(1:5) %>% 
  as.data.frame()

# Or convert the whole thing from UTF-8 to ASCII using base::iconv()
agencies %>% 
  mutate_all(funs(iconv(.,'utf-8', 'ascii', sub=''))) %>% 
  slice(1:5) %>% 
  as.data.frame()

```

### Resources

  + `stringr` [vignette](https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html) 
  + Stackoverflow question about removing unicode characters: [How to remove strange characters using gsub in R?](https://stackoverflow.com/q/38828620)
  + FileFormate.Info: [A to Z Index of Unicode Characters: starting with 'A'](https://www.fileformat.info/info/unicode/char/a.htm)